---
title: "EOS_Tantramar_WQ"
author: 
  - Miranda Corkum
  - Douglas Campbell 
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This .RMD is used to create exploratory plots of EOS Eco-Energy Inc. 2021 Tantramar River Watershed water quality monitoring data using a "DataStream" formatted dataset. Source data downloaded from Atlantic DataStream (https://atlanticdatastream.ca) on 8February2022.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up libraries (or 'packages') - these contain additions to base R.

```{r load libraries}
library(tidyverse)
#tidyverse set of packages for data wrangling and plotting; includes ggplot2
library(readr)
library(lubridate)
#assists with date formats
library(ggplot2)
library(dplyr)
library(ggbreak)

```

# Citations of software packages
```{r cite software packages}
citation()
citation("tidyverse")
citation("readr")
citation("lubridate")
citation("ggplot2")
citation(dplyr)
citation(ggbreak)

```

# Project Variables

Assign project-specific values to variables

```{r set project variables}
Project <- "EOS_Tantramar_WQ"

DataIn <- "DataIn"

ProcessData <- "ProcessData"

Plots <- "Plots"

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'
```

```{r list files in data folder}
DataFiles <- list.files(DataIn)
DataFiles
```

```{r set target file}
TargetFile <- "DATA.csv"
TargetFileTrim <- TargetFile %>%
  str_replace_all(pattern = " ", replacement =  "_")

TargetFileTrim
```

\#\#Import a tidied data object previously stored as .csv object

```{r read csv}
TidyData <- read_csv(file = file.path(DataIn, TargetFile, fsep =.Platform$file.sep))
TidyData[1:10, ]

```

```{r remove NA}
TidyData <- TidyData %>%
  discard(~all(is.na(.x))) %>%
  map_df(~.x)
```

\#\#Extract month from "ActivityStartDate" and create new column.

```{r add column month}
head(month(TidyData$ActivityStartDate))

TidyData2 <- TidyData %>%
  mutate(Month = month(ActivityStartDate, label = TRUE))

TidyData2[1:5, "Month"]
```

\#\#Subset rows of dataframe: want only data from 2021 for all plots below

```{r subset desired data range}
StartDate <- '2021-01-01'
TidyData3 <- filter(TidyData2, ActivityStartDate > StartDate)

TidyData3[1:5,]
```

\#\#Look at Column Names to check for non-standard names

```{r column names}
colnames(TidyData3)
```

\#Plot RPC bacteria data (*E. coli*) as a grouped bar graph

```{r plot E coli, fig.height=5, fig.width=7}
EcoliPlot <-
ggplot(subset(TidyData3, CharacteristicName == "Escherichia coli"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 400, linetype='dotted', col = 'red') +
  annotate("text", x = "Reservoir Brook", y = 400, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Escherichia coli count (MPN)")

EcoliPlot
```

\#Plot RPC water chemistry data (Total Phosphorus) as a grouped bar graph

```{r plot total P, fig.height=5, fig.width=7}
TotalPPlot <-
ggplot(subset(TidyData3, CharacteristicName == "Total Phosphorus, mixed forms"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_y_break(c(0.2, 0.5)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 0.03, linetype='dotted', col = 'red') +
  labs(x = "Monitoring Location", 
       y = "Total Phosphorus (mg/L)")

TotalPPlot
```

\#\#Subset rows of dataframe: want only in-situ probe data (some parameters were collected in-situ and also in the lab)

```{r subset in-situ data}
TidyData4 <- filter(TidyData3, ActivityType == "Field Msr/Obs-Portable Data Logger")

TidyData4[1:5,]
```

\#Monthly in-situ temperature plot (20 degC cut off) as a grouped bar graph

```{r plot insitu temp, fig.height=5, fig.width=7, echo=TRUE}
InsituWTempPlot <-
ggplot(subset(TidyData4, CharacteristicName == "Temperature, water"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 20, linetype='dotted', col = 'red') +
  labs(x = "Monitoring Location", 
       y = "Water Temperature (deg C)")

InsituWTempPlot
```

\#Monthly in-situ pH plot (acceptable long-term range 6.5 - 9) as a grouped bar graph

```{r plot insitu pH, fig.height=5, fig.width=7}
#adjust Y axis scale to reach 9
InsitupHPlot <-
ggplot(subset(TidyData4, CharacteristicName == "pH"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  coord_cartesian(ylim = c(5,9)) +
  labs(x = "Monitoring Location", 
       y = "pH")

InsitupHPlot
```

\#Monthly in-situ pH plot (acceptable long-term range 6.5 - 9) using facet wrap

```{r plot insitu pH facet wrap}
ggplot(data = subset(TidyData4, CharacteristicName == "pH"),
    aes(x = as.numeric(ActivityStartDate), y = ResultValue)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  theme_bw() +
  coord_cartesian(ylim = c(5,9)) +
  facet_wrap(facets = vars(MonitoringLocationName)) +
  labs(x = "Monitoring Location", 
       y = "pH")
```

\#Monthly in-situ DO plot (6.5 mg/L cut off) as a grouped bar graph

```{r plot insitu DO, fig.height=5, fig.width=7}
InsituDOPlot <-
ggplot(subset(TidyData4, CharacteristicName == "Dissolved oxygen (DO)"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  labs(x = "Monitoring Location", 
       y = "Dissolved oxygen (mg/L)")

InsituDOPlot
```

\#Monthly in-situ Total Dissolved Solids plot as a grouped bar graph

```{r plot insitu TDS, fig.height=5, fig.width=7}
InsituTDSPlot <-
ggplot(subset(TidyData4, CharacteristicName == "Total dissolved solids"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Monitoring Location", 
       y = "Total Dissolved Solids (ppm)")

InsituTDSPlot
```

\#Monthly in-situ Conductivity plot as a grouped bar graph

```{r plot insitu conductivity, fig.height=5, fig.width=7}
InsituCondPlot <-
ggplot(subset(TidyData4, CharacteristicName == "Conductivity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Monitoring Location", 
       y = "Conductivity (uS/cm)")

InsituCondPlot
```

\#Monthly in-situ Salinity plot as a grouped bar graph

```{r plot insitu salinity, fig.height=5, fig.width=7}
InsituSalPlot <-
ggplot(subset(TidyData4, CharacteristicName == "Salinity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Monitoring Location", 
       y = "Salinity (ppt)")

InsituSalPlot
```

\#Monthly AVG Chlorophyll-a plot (bottle sample measured at MtA ACME lab) as a grouped bar graph

```{r plot Chla, fig.height=5, fig.width=7}
ChlaPlot <-
ggplot(subset(TidyData3, CharacteristicName == "Chlorophyll a, corrected for pheophytin"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Monitoring Location", 
       y = "Chlorophyll-a (ug/L)")

ChlaPlot
```


```{r save all plots}
#EcoliPlot
#TotalPPlot
#InsituWTempPlot
#InsitupHPlot
#InsituDOPlot
#InsituTDSPlot
#InsituCondPlot
#InsituSalPlot
#ChlaPlot

ggsave(paste("EcoliPlot",".png",sep = ""), plot = EcoliPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("TotalPPlot",".png",sep = ""), plot = TotalPPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituWTempPlot",".png",sep = ""), plot = InsituWTempPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsitupHPlot",".png",sep = ""), plot = InsitupHPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituDOPlot",".png",sep = ""), plot = InsituDOPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituTDSPlot",".png",sep = ""), plot = InsituTDSPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituCondPlot",".png",sep = ""), plot = InsituCondPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituSalPlot",".png",sep = ""), plot = InsituSalPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("ChlaPlot",".png",sep = ""), plot = ChlaPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

```


