---
title: "EOS_CapeTormentine_WQ"
author: 
  - Miranda Corkum
  - Douglas Campbell 
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
csl: frontiers-in-marine-science.csl
bibliography: FreshWaterQuality
---

# Introduction

This .RMD is used to create exploratory plots of EOS Eco-Energy Inc. 2022 Cape Tormentine Peninsula Watershed Water Quality Monitoring Data. The 2022 dataset is accessed directly from DataStream using an API key.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages
Should DataStream R packages be installed here as well?
```{r install packages}
#install.packages("kableExtra")

```


# Set up libraries (or 'packages') 

These contain additions to base R
```{r load libraries}
#tidyverse set of packages for data wrangling and plotting; includes ggplot2 and readr
library(tidyverse)

#assists with date formats
library(lubridate)

library(kableExtra)

library(dplyr)

library(ggbreak)

```

# Project Variables

Assign project-specific values to variables

```{r set project variables}
Project <- "EOS_CapeTormentine_WQ"

DataIn <- file.path("..", "DataIn")

ProcessData <- "ProcessData"

TidiedData <- file.path("..", "TidiedData")

Figures <- file.path("..", "Figures")

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'
```

# Access data - DataStream **only run this code once**

For Tantramar 2021 dataset, data was obtained by downoading a .csv from DataStream, saving locally and importing into R.
Read in 2022 Cape Tormentine Peninsula Watershed data directly from Atlantic DataStream using API key.
Install the datastreamr package through Github to be able to read Atlantic Data Stream data into R.
```{r install datastreamr}
#devtools::install_github("gordonfn/datastreamr", force=TRUE)
library(datastreamr)
```


# Data Import **only run this code once**

Query "Cape Tormentine Peninsula Watershed Water Quality Monitoring" data set on Atlantic Data Stream.
*Patrick Sinclair - need to search for locations independently first, acquire Id, then search for records.
```{r DataStream data direct import}
#Mount Allison Phyto Lab API key: "GVy8yPkFk8oeAVZdYmRafcqfbVlTlq2Q" 
#EOS Cape Tormentine Peninsula Watershed Water Quality Monitoring data set DOI: "10.25976/af39-mv83"
#Read in test location "Gaspereau River Roundabout" (CT-001)

#ds_locations("GVy8yPkFk8oeAVZdYmRafcqfbVlTlq2Q",
             #filter=c("Name=Gaspereau River Roundabout", "DOI=10.25976/af39-mv83"))
            
TidyData <- ds_records("GVy8yPkFk8oeAVZdYmRafcqfbVlTlq2Q",
           filter=c("DOI=10.25976/af39-mv83", "ActivityStartDate>2022-01-01"),
           orderby = "ActivityStartDate", top=5000)

TidyData[1:10, ]
#View(TidyData)

#Attributation -> Attribution Licence (ODC-By) v1.0 "https://opendatacommons.org/licenses/by/1-0/"
#DOI -> https://doi.org/10.25976/af39-mv83
#License -> API
#Citation -> EOS Eco-Energy. 2022-09-30. "Cape Tormentine Peninsula Watershed Water Quality Monitoring" (dataset). 4.1.0. DataStream. https://doi.org/10.25976/af39-mv83.

```

## Save the imported data for further data exploration **only run this code once** 
(as a backup for DataStream direct access - can be problems with internet connection and/or datastream site issues)
.RDS is an internal R data format for rapid re-import into other RNotebooks or scripts
```{r save data}
saveRDS(TidyData, file = file.path(DataIn, paste("CapeTormentine", ".Rds", sep = ""), fsep = .Platform$file.sep))

```

*NO MONITORING LOCATION ID IN API ACCESSED DATASET*
## Load stored .Rds file
```{r load data}
 TidyData <- readRDS(DataIn, file = file.path(DataIn, paste("CapeTormentine", ".Rds", sep = ""), fsep = .Platform$file.sep))

```

**This file contains "MonitoringLocationID" but code below does not run properly?**
# Import a tidied data object previously stored as .csv object
```{r read csv}
TidyData <- read_csv(file = file.path(DataIn, "Cape_Tormentine_Peninsula_Watershed_Water_Quality_Monitoring_2022.csv", fsep =.Platform$file.sep))

TidyData[1:10, ]
#View(TidyData)

```

DC: Consider select(- remove columns which are all 'NA' Cite code retrieved from Google that is not part of Tidyverse and that we don't necessarily understand!Drops from 42 to 30 columns. TO DO: the AnalysisStartDate column is remaining because there are entries for July 11. Inconsistent - update on DataStream.

```{r remove NA columns}
#https://community.rstudio.com/t/drop-all-na-columns-from-a-dataframe/5844/6
TidyData <- TidyData %>%
  discard(~all(is.na(.x))) %>%
  map_df(~.x)
  #mutate(ActivityStartDate = ymd(ActivityStartDate)) 
  
#TidyData[1:10, ]
```

\#\#Extract month from "ActivityStartDate" and create new column (= 31 columns)

```{r add column month}
head(month(TidyData$ActivityStartDate))

TidyData <- TidyData %>%
  mutate(Month = month(ActivityStartDate, label = TRUE))

#View(TidyData)
#TidyData[1:5, "Month"]

#rm(TidyData)
#or, once running overwrite data objects onto same variable names
```

DC: give unique names to all chunks. RMarkdown commands in chunk headers control whether: -chunk appears -chunk output appears etc. upon knit 'echo' 'include'...

## Subset data - Remove Chapman Brook Route 955 and Scott Brook off Noonan Road
```{r subset data}
TidyData2 <- filter(TidyData, MonitoringLocationName != "Chapman Brook Route 955") %>%
      filter(MonitoringLocationName != "Scott Brook off Noonan Road")

#TidyData2[1:10,]

```


# Data Plots

## RPC bacteria data (*Escherichia coli*) as a grouped bar graph

DC: consider dynamic generation of graphic titles, subtitles and axes labels. Can be fussy to set up, possibly not worth the time, but can limit risk of generating mis-labelled graphics.
```{r plot E coli, fig.height=5, fig.width=7}
EcoliPlot <-
ggplot(subset(TidyData2, CharacteristicName == "Escherichia coli"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_break(c(2000, 5500)) +
  #scale_y_continuous(aes(breaks = seq(0, max(ResultValue), by = 0.1))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 400, linetype='dotted', col = 'red') +
  #annotate("text", x = "Blacklock Brook Route 955", y = 400, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Escherichia coli count (MPN)")

EcoliPlot

#DC prefers piping into a filter but this doesn't work here?
#TidyData3 %>%
  #filter(CharacteristicName == "Escherichia coli") %>%
  #ggplot() +
  #geom_bar(aes(x = MonitoringLocationName, y = ResultValue, fill = Month)
  
```

## RPC water chemistry data (Total Phosphorus) as a grouped bar graph

```{r plot total P, fig.height=5, fig.width=7}
TotalPPlot <-
ggplot(subset(TidyData2, CharacteristicName == "Total Phosphorus, mixed forms"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  #scale_y_break(c(0.2, 0.5)) +
  #scale_y_continuous(aes(breaks = seq(0, max(ResultValue), by = 0.1))) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 0.03, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 0.035, label = "", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Total Phosphorus (mg/L)")

TotalPPlot
```

\#\#Subset rows of dataframe: want only in-situ probe data (some parameters were collected in-situ and also in the lab)

```{r subset in-situ data}
InsituData <- filter(TidyData2, ActivityType == "Field Msr/Obs-Portable Data Logger")

InsituData[1:5,]
```

Naaman:
echo=TRUE - this will show graph in knit output
echo=FALSE - this will hide graph in knit output
include=TRUE - will show both code and graph in knit output
include=FALSE - will hide both code and graph

## Monthly in-situ temperature data (20 degC cut off) as a grouped bar graph

```{r plot insitu temp, fig.height=5, fig.width=7, echo=TRUE}
InsituWTempPlot <-
ggplot(subset(InsituData, CharacteristicName == "Temperature, water"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 20, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Water Temperature (deg C)")

InsituWTempPlot
```

## Monthly in-situ pH data (acceptable long-term range 6.5 - 9)

```{r plot insitu pH, fig.height=5, fig.width=7}
#adjust Y axis scale to reach 9
InsitupHPlot <-
ggplot(subset(InsituData, CharacteristicName == "pH"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  coord_cartesian(ylim = c(2,9)) +
  labs(x = "Monitoring Location", 
       y = "pH")

InsitupHPlot
```

# ```{r plot insitu pH facet wrap}
# ggplot(data = subset(InsituData, CharacteristicName == "pH"),
#     aes(x = as.numeric(ActivityStartDate), y = ResultValue)) + 
#   geom_point() +
#   geom_smooth(method = "lm") + 
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
#   theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
#   theme_bw() +
#   coord_cartesian(ylim = c(5,9)) +
#   facet_wrap(facets = vars(MonitoringLocationName)) +
#   labs(x = "Monitoring Location", 
#        y = "pH")
# ```

## Monthly in-situ DO plot (6.5 mg/L cut off)

```{r plot insitu DO, fig.height=5, fig.width=7}
InsituDOPlot <-
ggplot(subset(InsituData, CharacteristicName == "Dissolved oxygen (DO)"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Dissolved oxygen (mg/L)")

InsituDOPlot
```

## Monthly in-situ Total Dissolved Solids

```{r plot insitu TDS, fig.height=5, fig.width=7}
InsituTDSPlot <-
ggplot(subset(InsituData, CharacteristicName == "Total dissolved solids"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  #scale_y_break(c(4500, 10000)) +
  #scale_y_continuous(aes(breaks = seq(0, max(ResultValue), by = 0.1))) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Total Dissolved Solids (ppm)")

InsituTDSPlot
```

```{r plot insitu TDS FW, fig.height=5, fig.width=7}
InsituTDSPlotFW <-
ggplot(subset(InsituData, CharacteristicName == "Total dissolved solids"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  coord_cartesian(ylim = c(0,325)) +
  labs(x = "Monitoring Location", 
       y = "Total Dissolved Solids (ppm)")

InsituTDSPlotFW
```


## Monthly in-situ Conductivity

```{r plot insitu conductivity, fig.height=5, fig.width=7}
InsituCondPlot <-
ggplot(subset(InsituData, CharacteristicName == "Conductivity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Conductivity (uS/cm)")

InsituCondPlot
```

```{r plot insitu conductivity FW, fig.height=5, fig.width=7}
InsituCondPlotFW <-
ggplot(subset(InsituData, CharacteristicName == "Conductivity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  coord_cartesian(ylim = c(0,225)) +
  labs(x = "Monitoring Location", 
       y = "Conductivity (uS/cm)")

InsituCondPlotFW
```


## Monthly in-situ Salinity

```{r plot insitu salinity, fig.height=5, fig.width=7}
InsituSalPlot <-
ggplot(subset(InsituData
              , CharacteristicName == "Salinity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  labs(x = "Monitoring Location", 
       y = "Salinity (ppt)")

InsituSalPlot
```

```{r plot insitu salinity, fig.height=5, fig.width=7}
InsituSalPlotFW <-
ggplot(subset(InsituData
              , CharacteristicName == "Salinity"),
    aes(x = MonitoringLocationName, y = ResultValue, fill = Month)) + 
  geom_bar(stat="identity", width=0.7, position=position_dodge(width=0.8)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust=0.5, hjust=1)) +
  #geom_hline(yintercept = 6.5, linetype='dotted', col = 'red') +
  #annotate("text", x = "Reservoir Brook", y = 20, label = "Upper Limit Health Canada", vjust = -0.5) +
  coord_cartesian(ylim = c(0,0.25)) +
  labs(x = "Monitoring Location", 
       y = "Salinity (ppt)")

InsituSalPlotFW
```


## Save all plots
```{r save all plots}
#EcoliPlot
#TotalPPlot
#InsituWTempPlot
#InsitupHPlot
#InsituDOPlot
#InsituTDSPlot
#InsituCondPlot
#InsituSalPlot


ggsave(paste("EcoliPlot",".png",sep = ""), plot = EcoliPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("TotalPPlot",".png",sep = ""), plot = TotalPPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituWTempPlot",".png",sep = ""), plot = InsituWTempPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsitupHPlot",".png",sep = ""), plot = InsitupHPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituDOPlot",".png",sep = ""), plot = InsituDOPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituTDSPlot",".png",sep = ""), plot = InsituTDSPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituTDSPlotFW",".png",sep = ""), plot = InsituTDSPlotFW, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituCondPlot",".png",sep = ""), plot = InsituCondPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituCondPlotFW",".png",sep = ""), plot = InsituCondPlotFW, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituSalPlot",".png",sep = ""), plot = InsituSalPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("InsituSalPlotFW",".png",sep = ""), plot = InsituSalPlotFW, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

```






Concatenate 2 columns in R
USING BASE R:
dataf$MY <- paste(dataf$Month, dataf$Year)

USING STRINGR:
library(stringr) 
dataf$Combined <- str_c(dataf$A, '', dataf$B)

i.e.
library(stringr)
dataf$SnakeNSize <- str_c(dataf$Snake," ", dataf$Size)

*Highlight every second line:
dt %>%
  kbl() %>%
  kable_material(c("striped", "hover"))

*Change font size:
kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", font_size = 7)

*Change angle of column text labels:
kbl(dt) %>%
  kable_paper("striped", full_width = F) %>%
  row_spec(0, angle = -45)

*Conditional Formatting:
  use "cell_spec" with conditional logic
  

  

Creating Data Tables: 
tidyverse
pivot_wider
kable

https://tidyr.tidyverse.org/reference/pivot_wider.html

https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

- Conditional formatting

TRY
#options(kableExtra.auto_format = FALSE)
#https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#installation


SAVE JUST INCASE

SBWA: HCO3, Br, Ca, CO3, Cl, F, K, Mg, Na, NH3T, NH3_Un, NO2, NO3, NOX, SO4, TKN, TN, TOC, TP-L

# ```{r pivot data frame chemistry}
# ChemistryData <- BottleData %>%
#   pivot_wider(names_from = CharacteristicName, values_from = ResultValue) %>%
#   select(c("MonitoringLocationName", "ActivityStartDate", "Alkalinity, carbonate mg/l", "Ammonia mg/l", "Bicarbonate mg/l", "Bromine mg/l", "Calcium mg/l", "Calcium carbonate mg/l", "Chloride mg/l", "Color TCU", "Conductivity uS/cm", "Escherichia coli MPN", "Fluoride mg/l", "Hardness, carbonate mg/l", "Inorganic nitrogen (nitrate and nitrite) mg/l", "Langelier Saturation Index None", "Magnesium mg/l", "Nitrate mg/l", "Nitrite mg/l", "Organic carbon mg/l", "pH, lab None", "pH, saturated None", "Potassium mg/l", "Sulfate mg/l", "Total dissolved solids mg/l", "Total Nitrogen, mixed forms mg/l", "Total Phosphorus, mixed forms mg/l", "Turbidity NTU"))
# 
# #ChemistryData[1:10, ]
# View(ChemistryData)
# ```

Al, As, B, Ba, Cd, Co, Cr, Cu, Fe, Li, Mn, Mo, Ni, Pb, Rb, Sb, Sr, U, V, Zn