---
title: "EOS_Tantramar_Weather"
author: 
  - Miranda Corkum
  - Emma Sutherland
  - Douglas Campbell
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
bibliography: RPackageCitations.bib
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by EOS Eco-Energy New Brunswick Environmental Trust.
ES was supported by the New Brunswick Wildlife Trust Fund and a Cleantech Internship.  
DAC was supported by the Canada Research Chairs.

# Overview
This project uses packages to support direct import of Environment and Climate Change Canada weather data into R.

# Materials & Methods
## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Load packages
Preliminary work before running document. This involves installing packages and updating data.

```{r packages, include = FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(weathercan)
library(cowplot)
library(devtools)
library(ggplot2)
library(readr)
library(scales)

```

```{r Weathercan update}
#These packages may need to be installed to update Weathercan data.
#Update the station data to get the most recent data.

install.packages(c('lutz', 'sf'))

```

## Citations of software packages
```{r cite software packages}
citation()
citation("tidyverse")
citation("lubridate")
citation("stringr")
citation("broom")
citation("weathercan")
citation("cowplot")
citation("devtools")
citation("ggplot2")
citation("readr")
citation("scales")

```

## Set Project Variables
```{r set project variables}
Project <- "EOS_Tantramar_WQ"

DataIn <- "DataIn"

ProcessData <- "ProcessData"

Plots <- "Plots"

#These variables are required for ECCC weather data access
EnCanStation1 <- "NAPPAN AUTO"
Latitude <- 45.75
Longitude <- -64.49

Start <- 20180101000000
End <-Sys.time()

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'

```

## Search for Enviroment Canada weather station
```{r weather data station}
#search for StationNames containing the value of EnCanStation set above
StationNames <- weathercan::stations_search(EnCanStation1)
StationNames

```

## Download Environment Canada daily weather data for NAPPAN AUTO 2021
```{r weather import 2021 NAPPAN AUTO}
#Weather data available for years 2003 to 2021; call data from 2021/05/01 to 2021/10/31. 
NappanWeatherDay <- weathercan::weather_dl(station_ids = 42083, start = "2021-05-01", end = "2021-10-31", interval = "day")

NappanWeatherDay[1:5,]

#View(NappanWeatherDay)

```

## Set timezone
```{r set timezone}
#force_tz is a lubridate function to set a timezone
#for daily measurements in 2021
NappanWeatherDay <- NappanWeatherDay %>%
  mutate(date = ymd(date))

```

## Plot Tantramar River Watershed daily max wind gust 2021
```{r plot daily max wind gust, fig.height=5, fig.width=7}
TantramarMaxWindPlot <- NappanWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = spd_max_gust), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
    labs(x = "Date", 
       y = "Wind spd_max_gust (km/hr)")

TantramarMaxWindPlot

```

## Plot Tantramar River Watershed daily mean temperature 2021
```{r plot daily mean temp, fig.height=5, fig.width=7}
TantramarMeanTempPlot <- NappanWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = mean_temp), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Date", 
       y = "Mean Daily Temperature (degC)")

TantramarMeanTempPlot

```

##Plot Tantramar River Watershed total daily precipitation 2021
```{r plot daily precip, fig.height=5, fig.width=7}
TantramarPrecipPlot <- NappanWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = total_precip), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Date", 
       y = "Total Daily Precipitation (mm)")

TantramarPrecipPlot

```


```{r save all plots}
#TantramarMaxWindPlot
#TantramarMeanTempPlot
#TantramarPrecipPlot


ggsave(paste("TantramarMaxWindPlot",".png",sep = ""), plot = TantramarMaxWindPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)
ggsave(paste("TantramarMeanTempPlot",".png",sep = ""), plot = TantramarMeanTempPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)
ggsave(paste("TantramarPrecipPlot",".png",sep = ""), plot = TantramarPrecipPlot, device = NULL, path = Plots,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

```
