---
title: "Tormentine_Weather"
author: 
- Miranda Corkum ^1^
- Sarah Arthur^1^
- Emma Sutherland^1^ 
- Douglas Campbell^1^*
date: "`r format(Sys.Date())`"
output: bookdown::html_document2
output: html_document
bibliography: RPackageCitations.bib
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by
SA was supported by
ES was supported by the New Brunswick Wildlife Trust Fund and a Cleantech Internship.  
DAC was supported by the Canada Research Chairs.


# Overview
This project uses packages to support direct import of Environment and Climate Change Canada weather data into R.

# What does this stations_meta() line do?
```{r introduction, child = "WaterWeatherIntro.Rmd"}
#stations_meta()
```

# Materials & Methods
## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
Preliminary work before running document. This involves installing packages and updating data.

```{r packages, include = FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(weathercan)
library(cowplot)
library(devtools)
library(scales)

#DC: Do not reload packages that are already included in base R - can cause version conflicts
#library(stringr)
#library(ggplot2)
#library(readr)

```


#These packages may need to be installed to update Weathercan data - *only run occasionally*
```{r Weathercan update}
install.packages(c('lutz', 'sf'))
#stations_dl()

```

## Citations of software packages

## Set Project Variables for site 1 and 2  
```{r set project variables}
Project <- "EOS_CapeTormentine_WQ"

DataIn <- "DataIn"

ProcessData <- "ProcessData"

Figures <- file.path("..", "Figures")

#These variables are required for ECCC weather data access
EnCanStation1 <-"MONCTON"
Latitude2 <- 46.11
Longitude2 <- 64.68

Start <- 20100101000000
End <-Sys.time()
#TrimStart <- 20200401000000
#TrimEnd <- 20200501000000
#TideTolerance <- 5

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'
```


## Search for Enviroment Canada weather station
```{r weather data station}
#search for StationNames containing the value of EnCanStation set above
StationNames <- weathercan::stations_search(EnCanStation1)
StationNames

#alternatively 
#stations_search("NAPPAN AUTO", interval = "day")

```


## Download Environment Canada daily weather data for "MONCTON / GREATER MONCTON ROMEO LEBLANC INTL A"
```{r weather import 2021}
#be careful; time is likely reported as AST without ADT 
# Station ID 54282 also has daily data for this year but NO precip data
#Call data from 2022/05/01 to 2022/10/31. 
MonctonWeatherDay <- weathercan::weather_dl(station_ids = 50309, start = "2022-05-01", end = "2022-10-31", interval = "day")

MonctonWeatherDay[1:5,]

#NappanWeatherMonth <- weathercan::weather_dl(station_ids = 42083, start = "2021-05-01", end = "2021-10-31", interval = "month")
#NappanWeatherMonth[1:5,]


#View(NappanWeatherDay)

```


```{r weather import 2018}

#NappanWeatherDay2018 <- weathercan::weather_dl(station_ids = 42083, start = "2018-05-01", end = "2018-10-31", interval = "day")

#NappanWeatherDay2018[1:5,]

#View(NappanWeatherDay2018)
```


## Set timezone
```{r set timezone}
#force_tz is a lubridate function to set a timezone
#for hourly measurements example
#KouchibouguacWeather <- KouchibouguacWeather %>%
  #mutate(time = force_tz(ymd_hms(time),"America/Halifax"))

#for daily measurements in 2021
MonctonWeatherDay <- MonctonWeatherDay %>%
  mutate(date = ymd(date))
#  mutate(date = as_date(force_tz(ymd(date),"America/Halifax")))

#for daily measurements in 2018
#NappanWeatherDay2018 <- NappanWeatherDay2018 %>%
#  mutate(date = force_tz(ymd(date),"America/Halifax"))

#for monthly measurements
#NappanWeatherMonth <- NappanWeatherMonth %>%
#  mutate(date = force_tz(ymd(date),"America/Halifax"))

```


##Plot Tantramar River Watershed daily max wind gust 2021
```{r plot daily max wind gust, fig.height=5, fig.width=7}
TormentineMaxWindPlot <- MonctonWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = spd_max_gust), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
    labs(x = "Date", 
       y = "Wind spd_max_gust (km/hr)")

TormentineMaxWindPlot
```

##Plot Tantramar River Watershed daily mean temperature 2021
Comment - Ask DC: Figure out how to 1) add all months to x axis labels; 2) overlay 2018 data? 3) indicate sampling day? Should April data be included? 
```{r plot daily mean temp, fig.height=5, fig.width=7}

TormentineMeanTempPlot <- MonctonWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = mean_temp), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Date", 
       y = "Mean Daily Temperature (degC)")

TormentineMeanTempPlot

```

##Plot Tantramar River Watershed total daily precipitation 2021
```{r plot daily precip, fig.height=5, fig.width=7}
TormentinePrecipPlot <- MonctonWeatherDay %>% 
  ggplot() +
  scale_x_date(date_breaks = "10 days", date_minor_breaks = "10 day", labels = date_format("%b-%d")) +
  geom_line(aes(x = date, y = total_precip), size = 0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
  labs(x = "Date", 
       y = "Total Daily Precipitation (mm)")

TormentinePrecipPlot
```


```{r save all plots}
#TantramarMaxWindPlot
#TantramarMeanTempPlot
#TantramarPrecipPlot


ggsave(filename = paste("TormentineMaxWindPlot",".png",sep = ""), plot = TormentineMaxWindPlot, device = NULL, path = Figures, scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)
ggsave(paste("TormentineMeanTempPlot",".png",sep = ""), plot = TormentineMeanTempPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)
ggsave(paste("TormentinePrecipPlot",".png",sep = ""), plot = TormentinePrecipPlot, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

```





```{r}
 
# 1. The x variable should be numeric in order to properly draw the line
#NappanWeatherDay$date <- as.numeric(NappanWeatherDay$date)
# 2. The y variable should also be numeric for the same reason
NappanWeatherDay$mean_temp <- as.numeric(NappanWeatherDay$mean_temp)
# 3. The variable used to color and split the data should be a factor so lines are properly drawn
#NappanWeatherDay$year <- factor(NappanWeatherDay$year)

TantramarTemp2 <- NappanWeatherDay %>% 
  ggplot() +
  geom_line(aes(x = date, y = mean_temp, color = year),size = 0.2) +
  #scale_x_continuous(breaks = 1:12) +
  #scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") 
  theme_bw() +
  labs(x = "Date", 
       y = "Mean Daily Temperature degC", 
       title = "Tantramar Mean Daily Temperature")

TantramarTemp2
```
